---
output: 
  html_document
---

```{r, message=FALSE}

library(here)
library(tibble)
library(ranger)
library(caret)
library(ggplot2)
library(pROC)

source(here("v2","src","evidencegeom.R"))

library(stringr)
library(survival)
library(survminer)
library(timeROC)
library(rms)
library(dplyr)

set.seed(42)


```

## MIMIC-IV : ICU mortality 24h feature aggregation

This dataset contains 24h feature aggregates of multiple measurements from the mimiciv_3_1_derived.vitalsign, chemistry, complete_blood_count, and bg tables. The following aggregations were computed during data extraction: # of measurements, measurement missing indicator, min, max, mean, median, IQR, early mean (before 12h), late mean (after 12h), # occurrences below safe lower bound for measurement, # occurrences above safe upper bound for measurement, # in low/med/high bins. In this notebook, we also compute measurement entropy using counts in low/med/high bins, fraction of occurrences below safe lower bound, and fraction of occurrences above safe upper bound.



```{r}

mimiciv_import <- read.csv(here("v2","src","data","mimiciv","mimiciv_ext_24h_features_wide_2.csv"))

```


```{r}

mimiciv_import$mortality <- as.factor(mimiciv_import$mortality)

```

```{r}

mimiciv_import$gender <- as.factor(mimiciv_import$gender)

```


```{r}



# -----------------------------------
# 1) entropy from *_bin_low/_normal/_high
# 2) *_missing -> factor
# 3) *_low_count / *_high_count -> fractions
# -----------------------------------

safe_entropy3 <- function(low, mid, high, eps = 1e-12) {
  total <- low + mid + high
  
  out <- rep(NA_real_, length(total))
  ok <- is.finite(total) & total > 0
  
  if (any(ok)) {
    p1 <- low[ok]  / total[ok]
    p2 <- mid[ok]  / total[ok]
    p3 <- high[ok] / total[ok]
    
    P <- cbind(p1, p2, p3)
    P[P <= 0 | !is.finite(P)] <- eps
    
    out[ok] <- -rowSums(P * log(P))
  }
  
  out
}

add_entropy_missing_frac <- function(df) {
  
  nms <- names(df)
  
  # -----------------------------
  # 1) entropy from bin counts
  # expects columns like:
  #   heart_rate_bin_low
  #   heart_rate_bin_normal
  #   heart_rate_bin_high
  # -----------------------------
  bin_low_cols <- nms[str_detect(nms, "_bin_low$")]
  
  for (low_col in bin_low_cols) {
    stem <- str_remove(low_col, "_bin_low$")
    mid_col  <- paste0(stem, "_bin_normal")
    high_col <- paste0(stem, "_bin_high")
    
    if (mid_col %in% names(df) && high_col %in% names(df)) {
      ent_col <- paste0(stem, "_entropy")
      df[[ent_col]] <- safe_entropy3(
        low  = df[[low_col]],
        mid  = df[[mid_col]],
        high = df[[high_col]]
      )
    }
  }
  
  # -----------------------------
  # 2) convert *_missing to factor
  # -----------------------------
  missing_cols <- nms[str_detect(nms, "_missing$")]
  
  if (length(missing_cols) > 0) {
    df <- df %>%
      mutate(across(all_of(missing_cols), ~ factor(.x, levels = c(0, 1))))
  }
  
  # -----------------------------
  # 3) low/high count fractions
  # denominator = corresponding *_count
  # examples:
  #   heart_rate_low_count / heart_rate_count
  #   heart_rate_high_count / heart_rate_count
  # -----------------------------
  low_count_cols  <- names(df)[str_detect(names(df), "_low_count$")]
  high_count_cols <- names(df)[str_detect(names(df), "_high_count$")]
  
  make_frac <- function(count_vec, denom_vec) {
    out <- rep(NA_real_, length(count_vec))
    ok <- is.finite(denom_vec) & denom_vec > 0 & is.finite(count_vec)
    out[ok] <- count_vec[ok] / denom_vec[ok]
    out
  }
  
  for (col in low_count_cols) {
    stem <- str_remove(col, "_low_count$")
    denom_col <- paste0(stem, "_count")
    frac_col  <- paste0(stem, "_low_frac")
    
    if (denom_col %in% names(df)) {
      df[[frac_col]] <- make_frac(df[[col]], df[[denom_col]])
    }
  }
  
  for (col in high_count_cols) {
    stem <- str_remove(col, "_high_count$")
    denom_col <- paste0(stem, "_count")
    frac_col  <- paste0(stem, "_high_frac")
    
    if (denom_col %in% names(df)) {
      df[[frac_col]] <- make_frac(df[[col]], df[[denom_col]])
    }
  }
  
  df
}

```



```{r}

mimiciv_import <- add_entropy_missing_frac(mimiciv_import)

```

```{r}

drop_bin_cols <- names(mimiciv_import)[str_detect(names(mimiciv_import), "_bin_(low|normal|high)$")]
mimic_clean <- mimiciv_import %>% dplyr::select(-all_of(drop_bin_cols))

```


```{r}

mimic_clean <- mimic_clean%>%select(-all_of(c("spo2_high_count","spo2_high_frac")))

```


## Train-Val-Test Split

```{r}

mimic_clean_train_idx <- caret::createDataPartition(mimic_clean$mortality, p=0.5, list=FALSE)
mimic_clean_train <- mimic_clean[mimic_clean_train_idx, ]
mimic_clean_eval <- mimic_clean[-mimic_clean_train_idx, ]

mimic_clean_val_idx <- createDataPartition(mimic_clean_eval$mortality, p=0.5, list=FALSE)
mimic_clean_val <- mimic_clean_eval[mimic_clean_val_idx, ]
mimic_clean_test <- mimic_clean_eval[-mimic_clean_val_idx, ]

```


## Evidence Geometry - Validation Set

### Fit Marginal Likelihood models on Training set

```{r}

alpha <- (nrow(mimic_clean_train[mimic_clean_train$mortality=="1", ])/nrow(mimic_clean_train))

```


```{r}

spec <- risk_spec(
   y_col="mortality", positive="1",
   features = setdiff(names(mimic_clean_train), c("mortality", "stay_id")),
   alpha = alpha,
   laplace = 1, ridge = 1e-3, winsor_p=0.01,
   weights=FALSE,
   weight_method = "mutual_info",
   numeric_candidates = c("gaussian", "lognormal", "gamma"),
   count_candidates = c("poisson", "negbinom"),
   fraction_candidates = c("logit_gaussian"),
   fraction_eps = 1e-3,
   numeric_val_frac = 0.2,
   numeric_min_n = 25,
   llr_cap_quantile = 0.01,
   mi_nbins = 10
 )

obj <- fit(mimic_clean_train%>%dplyr::select(-stay_id), spec, k_eigen=4, k_energy=4, energy_ref="both")


```

### Compute Log-Likelihood Ratio (LLR) Matrix for Validation Set. Compute geometric risk signals

```{r}

Lval <- loglik_matrices(mimic_clean_val%>%dplyr::select(-stay_id), obj$fit, alpha=spec$alpha)
# val/test scoring
scores_val <- bind_cols(stay_id=mimic_clean_val$stay_id, 
                            mortality=mimic_clean_val$mortality, 
                            score_risk(Lval$l_pos, Lval$l_neg,  obj$fit$weights, obj$geom, 
                                       spec$alpha, spec$eps))


```

### Train Random Forest on Train Set

```{r}

rf_train_1 <- ranger(formula = mortality ~ ., data=mimic_clean_train%>%dplyr::select(-stay_id),
                         mtry=10, min.node.size=7, num.trees=500, probability = TRUE,
                         keep.inbag = TRUE)

```

### Append RF predictions on Validation set to geometric risk dataframe

```{r}

scores_val <- bind_cols(scores_val,
                            p = predict(rf_train_1, mimic_clean_val%>%dplyr::select(-stay_id, -mortality))$predictions[,"1"])

```


### Check assigned marginal likelihood families and compare parameters against population statistics

```{r}


feature_sanity_check <- function(df_val,
                                 L_val,
                                 y_col,
                                 positive_label = "1",
                                 feature_type_summary = NULL) {
  y <- as.character(df_val[[y_col]])
  feats <- intersect(colnames(L_val), setdiff(names(df_val), y_col))
  
  get_auc <- function(response, predictor) {
    ok <- is.finite(predictor) & !is.na(response)
    if (sum(ok) < 20) return(NA_real_)
    rr <- tryCatch(
      pROC::roc(response = response[ok], predictor = predictor[ok], quiet = TRUE),
      error = function(e) NULL
    )
    if (is.null(rr)) return(NA_real_)
    as.numeric(rr$auc)
  }
  
  safe_mode <- function(x) {
    x <- x[!is.na(x)]
    if (length(x) == 0) return(NA_character_)
    tab <- sort(table(x), decreasing = TRUE)
    names(tab)[1]
  }
  
  out <- lapply(feats, function(f) {
    x <- df_val[[f]]
    l <- L_val[, f]
    
    pos_idx <- y == positive_label
    neg_idx <- y != positive_label
    
    is_num <- is.numeric(x) || is.integer(x)
    
    if (is_num) {
      raw_mean_pos   <- mean(x[pos_idx], na.rm = TRUE)
      raw_mean_neg   <- mean(x[neg_idx], na.rm = TRUE)
      raw_median_pos <- median(x[pos_idx], na.rm = TRUE)
      raw_median_neg <- median(x[neg_idx], na.rm = TRUE)
      raw_auc        <- get_auc(y, as.numeric(x))
      raw_dir        <- sign(raw_mean_pos - raw_mean_neg)
      raw_mode_pos   <- NA_character_
      raw_mode_neg   <- NA_character_
    } else {
      raw_mean_pos   <- NA_real_
      raw_mean_neg   <- NA_real_
      raw_median_pos <- NA_real_
      raw_median_neg <- NA_real_
      raw_auc        <- NA_real_
      raw_dir        <- NA_real_
      raw_mode_pos   <- safe_mode(as.character(x[pos_idx]))
      raw_mode_neg   <- safe_mode(as.character(x[neg_idx]))
    }
    
    llr_mean_pos <- mean(l[pos_idx], na.rm = TRUE)
    llr_mean_neg <- mean(l[neg_idx], na.rm = TRUE)
    llr_auc      <- get_auc(y, as.numeric(l))
    llr_dir      <- sign(llr_mean_pos - llr_mean_neg)
    
    data.frame(
      feature = f,
      raw_type = class(x)[1],
      raw_mean_pos = raw_mean_pos,
      raw_mean_neg = raw_mean_neg,
      raw_median_pos = raw_median_pos,
      raw_median_neg = raw_median_neg,
      raw_mode_pos = raw_mode_pos,
      raw_mode_neg = raw_mode_neg,
      raw_auc = raw_auc,
      llr_mean_pos = llr_mean_pos,
      llr_mean_neg = llr_mean_neg,
      llr_auc = llr_auc,
      raw_dir = raw_dir,
      llr_dir = llr_dir,
      direction_agrees = ifelse(is.na(raw_dir) | is.na(llr_dir), NA, raw_dir == llr_dir),
      llr_minus_raw_auc = llr_auc - raw_auc,
      stringsAsFactors = FALSE
    )
  })
  
  out <- bind_rows(out)
  
  if (!is.null(feature_type_summary)) {
    out <- left_join(out, feature_type_summary, by = "feature")
  }
  
  out %>%
    mutate(
      suspicious = case_when(
        !is.na(direction_agrees) & !direction_agrees ~ "direction_mismatch",
        !is.na(raw_auc) & !is.na(llr_auc) & raw_auc < 0.55 & llr_auc > 0.75 ~ "llr_much_stronger_than_raw",
        !is.na(raw_auc) & !is.na(llr_auc) & raw_auc > 0.70 & llr_auc < 0.60 ~ "llr_weaker_than_raw",
        TRUE ~ ""
      )
    ) %>%
    arrange(desc(llr_auc))
}
```

```{r}

sanity_tbl <- feature_sanity_check(mimic_clean_val, L_val=Lval$L, y_col="mortality", positive_label = "1", feature_type_summary = NULL)

```

```{r}

likelihood_tbl <- print_feature_likelihoods(obj)

```


```{r}

check_tbl <- left_join(likelihood_tbl, sanity_tbl, by=join_by(feature))%>%
  dplyr::select(feature, raw_auc, llr_auc, class, likelihood_family, raw_median_pos, raw_median_neg, parameters)

```

```{r}


check_tbl

```


### Basic Plots

1. Density of sum of marginal LLRs versus class  
2. Density of d_dist versus class  
3. Density of proj versus class  
4. RF p versus sum of marginal LLRs by class  
5. Logit transform of RF p versus sum of marginal LLRs  
6. RF p versus d_dist by class  
7. RF p versus proj by class  
8. proj versus d_dist by class  
9. proj versus d_dist in -4 < d_dist < 4 region  

```{r}

ggplot(scores_val) +
  geom_density(aes(x=p, fill=mortality), alpha=0.6) +
  xlab("Random Forest probability of mortality") +
  theme_minimal()

ggplot(scores_val) +
  geom_density(aes(x=l, fill=mortality), alpha=0.6) +
  xlab("Sum of relative marginal evidence from all features") +
  theme_minimal()

ggplot(scores_val) +
  geom_density(aes(x=d_dist, fill=mortality), alpha=0.6) +
  xlab("d_dist") +
  theme_minimal()

ggplot(scores_val) +
  geom_density(aes(x=proj, fill=mortality), alpha=0.6) +
  xlab("proj") +
  theme_minimal()

ggplot(scores_val) +
  geom_point(aes(x=l, y=p, color=mortality), alpha=0.4) +
  ylab("Random forest mortality probability (p)") +
  theme_minimal()

ggplot(scores_val) +
  geom_point(aes(x=qlogis(p), y=l, color=mortality), alpha=0.4) +
  xlab("log-odds of random forest mortality probability (p)") +
  theme_minimal()

ggplot(scores_val) +
  geom_point(aes(x=p, y=d_dist, color=mortality), alpha=0.4) +
  xlab("Random forest mortality probability (p)") +
  theme_minimal()

ggplot(scores_val) +
  geom_point(aes(x=p, y=proj, color=mortality), alpha=0.4) +
  xlab("Random forest mortality probability (p)") +
  theme_minimal()

ggplot(scores_val) +
  geom_point(aes(x=d_dist, y=proj, color=mortality), alpha=0.4) +
  theme_minimal()

ggplot(scores_val%>%filter(d_dist > -4, d_dist < 4)) +
  geom_point(aes(x=d_dist, y=proj, color=mortality), alpha=0.4) +
  theme_minimal()




```


### Count and Mortality Rate Table for d_dist=0, proj=0

```{r}

scores_val %>%
  mutate(region = case_when(
    proj <=0 & d_dist<0 ~ "A",
    proj <=0 & d_dist>=0 ~ "B",
    proj >0 & d_dist<0 ~ "C",
    proj >0 & d_dist>=0 ~ "D"
  )) %>%
  group_by(region) %>%
  summarise(
    mortality_rate = mean(mortality=="1"),
    n=n()
  )

```

### Mortality rate in d_dist < 0 region

```{r}

val_lowrisk <- scores_val%>%filter(d_dist < 0)

```

```{r}

nrow(val_lowrisk%>%filter(mortality=="1")) / nrow(val_lowrisk)

```


### Survived cases in d_dist < 0 as a fraction of total survived cases

```{r}

nrow(val_lowrisk%>%filter(mortality=="0")) / nrow(scores_val%>%filter(mortality=="0"))

```

### Correlation between RF p, d_dist, proj

```{r}

cor(scores_val$p, scores_val$d_dist)
cor(scores_val$p, scores_val$proj)

```


### AUC-ROC of RF p, d_dist, proj

```{r}


pROC::roc(scores_val$mortality, scores_val$p)

pROC::roc(scores_val$mortality, scores_val$d_dist)

pROC::roc(scores_val$mortality, scores_val$proj)

```

### Mortality Rate and case counts for binned d_dist

```{r}

scores_val %>% filter(d_dist >= 0, d_dist < 20) %>% mutate(d_dist_bin = cut(d_dist, breaks = seq(0, 20, by = 1))) %>% group_by(d_dist_bin) %>% dplyr::summarize(count = n(), mortality_rate = mean(mortality == "1"))


scores_val %>% filter(d_dist >= 0, d_dist < 20) %>% mutate(d_dist_bin = cut(d_dist, breaks = seq(0, 20, by = 1))) %>%
  group_by(d_dist_bin) %>%
  summarise(
    mortality_rate = mean(mortality=="1")
  ) %>% ggplot(aes(x=d_dist_bin, y=mortality_rate, group = 1)) + geom_point() + geom_line() + theme_minimal()


```

### Cases with $d_{dist} \geq 2$ as a fraction of all cases with non-negative d_dist

```{r}

nrow(scores_val%>%filter(d_dist >= 2)) / nrow(scores_val%>%filter(d_dist >= 0))

```

### RF p vs mortality rate

```{r}

scores_val %>% mutate(p_bin = cut(p, breaks = 25)) %>% group_by(p_bin) %>% dplyr::summarize(count = n(), mortality_rate = mean(mortality == "1"))


scores_val %>% mutate(p_bin = cut(p, breaks = 25)) %>%
  group_by(p_bin) %>%
  summarise(
    mortality_rate = mean(mortality=="1")
  ) %>% ggplot(aes(x=p_bin, y=mortality_rate, group = 1)) + geom_point() + geom_line() + theme_minimal()


```


### Obtain LLR matrix and geometric risk signals for Training Set to calculate **marginal median and IQR for d_dist < 0 region**

```{r}

Ltrain <- loglik_matrices(mimic_clean_train%>%dplyr::select(-stay_id), obj$fit, alpha=spec$alpha)

scores_train <- bind_cols(stay_id=mimic_clean_train$stay_id, 
                            mortality=mimic_clean_train$mortality, 
                            score_risk(Ltrain$l_pos, Ltrain$l_neg,  obj$fit$weights, obj$geom, 
                                       spec$alpha, spec$eps))

```


### Calculate Positive LLR sums - overall and feature group-wise

All aggregated statistics for every measurement are group together as one feature group.  

#### Triage Rules  

**1. Primary Triage using $d_{dist}$**  

$d_{dist} < 0$ -> **Low Risk**  
$0 \leq d_{dist} < 2$ -> **Ambiguous/Moderate Risk**  
$d_{dist} \geq 2$ -> **High Risk**  
  
**2. Positive Evidence Scoring**  
  
$S_i = \sum_{j=1}^{p} \max(LLR_{ij}, 0)$  

$S_{\text{stable}} = \{ S_i \mid d_{dist,i} < 0 \}$  
  
$z_i = \frac{S_i - \operatorname{Median}(S_{\text{stable}})}{\operatorname{IQR}(S_{\text{stable}})\sqrt{p}}$  

In words, $z$ = Positive LLR sum relative to $d_{dist} < 0$ region scaled by sqrt(# features)  
  
$z < 0$ -> **Weak Positive Evidence**  
$0 \leq z < 2$ -> **Moderate Positive Evidence**  
$z \geq 2$ -> **Strong Positive Evidence**  
  
**3. Feature group contributions to positive evidence**  
  
Calculate feature group-wise raw positive LLR sums and compute **percentage** over total raw positive LLR sum  
  
```{r}

extract_signal_family <- function(feature_names) {
  feature_names %>%
    str_remove("_(mean|median|max|min|iqr|sd|early_mean|late_mean)$") %>%
    str_remove("_(count|missing|low_count|high_count|low_frac|high_frac)$") %>%
    str_remove("_(bin_low|bin_normal|bin_high)$") %>%
    str_remove("_entropy$")
}

extract_L_matrix <- function(x) {
  if (is.matrix(x) || is.data.frame(x)) {
    return(as.matrix(x))
  }
  if (is.list(x) && !is.null(x$L)) {
    return(as.matrix(x$L))
  }
  stop("Input must be either an LLR matrix or a loglik_matrices() object containing $L.")
}

# ------------------------------------------------------------
# Compute overall positive evidence and family-level evidence
# ------------------------------------------------------------
compute_positive_llr_scores <- function(L,
                                        normalize_family = TRUE,
                                        normalize_overall = TRUE,
                                        family_map = NULL,
                                        eps = 1e-12) {
  L <- as.matrix(L)
  Lpos <- pmax(L, 0)
  
  p <- ncol(Lpos)
  
  # overall positive evidence
  pos_sum_raw <- rowSums(Lpos, na.rm = TRUE)
  pos_sum <- if (normalize_overall) pos_sum_raw / sqrt(p) else pos_sum_raw
  
  # family map
  if (is.null(family_map)) {
    fam <- extract_signal_family(colnames(L))
    family_map <- split(colnames(L), fam)
  } else {
    needed <- unique(unlist(family_map, use.names = FALSE))
    missing <- setdiff(needed, colnames(L))
    if (length(missing) > 0) {
      stop("L is missing features required by family_map: ",
           paste(missing, collapse = ", "))
    }
  }
  
  family_map <- family_map[sort(names(family_map))]
  
  # raw and normalized family evidence
  fam_raw <- lapply(names(family_map), function(g) {
    feats <- family_map[[g]]
    X <- Lpos[, feats, drop = FALSE]
    rowSums(X, na.rm = TRUE)
  })
  names(fam_raw) <- names(family_map)
  fam_raw <- as.data.frame(fam_raw, check.names = FALSE)
  names(fam_raw) <- paste0(names(fam_raw), "_axis_raw")
  
  fam_scores <- lapply(names(family_map), function(g) {
    feats <- family_map[[g]]
    X <- Lpos[, feats, drop = FALSE]
    s <- rowSums(X, na.rm = TRUE)
    if (normalize_family) {
      s <- s / sqrt(length(feats))
    }
    s
  })
  names(fam_scores) <- names(family_map)
  fam_scores <- as.data.frame(fam_scores, check.names = FALSE)
  names(fam_scores) <- paste0(names(fam_scores), "_axis")
  
  # percentage contribution using RAW family sums / RAW total positive sum
  fam_pct <- fam_raw
  for (j in seq_len(ncol(fam_pct))) {
    fam_pct[[j]] <- ifelse(pos_sum_raw > eps, fam_raw[[j]] / pos_sum_raw, 0)
  }
  names(fam_pct) <- paste0(gsub("_axis_raw$", "", names(fam_raw)), "_pct")
  
  list(
    pos_sum_raw = pos_sum_raw,
    pos_sum = pos_sum,
    fam_scores = fam_scores,
    fam_raw = fam_raw,
    fam_pct = fam_pct,
    family_map = family_map
  )
}

# ------------------------------------------------------------
# Stable scaling for overall positive evidence
# ------------------------------------------------------------
make_stable_scaler <- function(x, min_iqr = 1e-4) {
  x <- x[is.finite(x)]
  
  med <- median(x, na.rm = TRUE)
  iqr <- IQR(x, na.rm = TRUE)
  iqr_ok <- is.finite(iqr) && iqr >= min_iqr
  
  list(
    median = med,
    iqr = iqr,
    iqr_ok = iqr_ok
  )
}

apply_stable_scaler <- function(x, scaler) {
  data.frame(
    raw = x,
    stable_z = if (scaler$iqr_ok) {
      (x - scaler$median) / scaler$iqr
    } else {
      x - scaler$median
    },
    stringsAsFactors = FALSE
  )
}

# ------------------------------------------------------------
# Main appender
# ------------------------------------------------------------
append_interpretability_scores <- function(scores_train,
                                           scores_val,
                                           ll_train,
                                           ll_val,
                                           stable_idx = NULL,
                                           normalize_family = TRUE,
                                           normalize_overall = TRUE,
                                           min_iqr = 1e-4) {
  if (is.null(stable_idx)) {
    stable_idx <- scores_train$d_dist < 0
  }
  
  if (!is.logical(stable_idx) || length(stable_idx) != nrow(scores_train)) {
    stop("stable_idx must be logical and align with scores_train rows.")
  }
  
  L_train <- extract_L_matrix(ll_train)
  L_val   <- extract_L_matrix(ll_val)
  
  if (nrow(L_train) != nrow(scores_train)) {
    stop("nrow(L_train) must match nrow(scores_train)")
  }
  if (nrow(L_val) != nrow(scores_val)) {
    stop("nrow(L_val) must match nrow(scores_val)")
  }
  
  # training-defined family map
  train_feats <- colnames(L_train)
  val_feats   <- colnames(L_val)
  
  missing_in_val <- setdiff(train_feats, val_feats)
  if (length(missing_in_val) > 0) {
    stop("Validation L matrix is missing training features: ",
         paste(missing_in_val, collapse = ", "))
  }
  
  L_val <- L_val[, train_feats, drop = FALSE]
  
  train_fam <- extract_signal_family(train_feats)
  family_map <- split(train_feats, train_fam)
  family_map <- family_map[sort(names(family_map))]
  
  train_pos <- compute_positive_llr_scores(
    L_train,
    normalize_family = normalize_family,
    normalize_overall = normalize_overall,
    family_map = family_map
  )
  
  val_pos <- compute_positive_llr_scores(
    L_val,
    normalize_family = normalize_family,
    normalize_overall = normalize_overall,
    family_map = family_map
  )
  
  # overall positive evidence
  scores_train$pos_sum_raw <- train_pos$pos_sum_raw
  scores_val$pos_sum_raw   <- val_pos$pos_sum_raw
  
  scores_train$pos_sum <- train_pos$pos_sum
  scores_val$pos_sum   <- val_pos$pos_sum
  
  pos_scaler <- make_stable_scaler(scores_train$pos_sum[stable_idx], min_iqr = min_iqr)
  pos_scaled_val <- apply_stable_scaler(scores_val$pos_sum, pos_scaler)
  scores_val$pos_sum_stable_z <- pos_scaled_val$stable_z
  
  # family scores and family percentages
  fam_cols <- names(train_pos$fam_scores)
  pct_cols <- names(train_pos$fam_pct)
  
  if (!setequal(fam_cols, names(val_pos$fam_scores))) {
    stop("Family axis names differ between training and validation.")
  }
  val_pos$fam_scores <- val_pos$fam_scores[, fam_cols, drop = FALSE]
  
  if (!setequal(pct_cols, names(val_pos$fam_pct))) {
    stop("Family percentage names differ between training and validation.")
  }
  val_pos$fam_pct <- val_pos$fam_pct[, pct_cols, drop = FALSE]
  
  family_scaler_info <- list()
  
  for (col in fam_cols) {
    train_vec <- train_pos$fam_scores[[col]]
    val_vec   <- val_pos$fam_scores[[col]]
    
    scores_train[[col]] <- train_vec
    scores_val[[col]]   <- val_vec
    
    sc <- make_stable_scaler(train_vec[stable_idx], min_iqr = min_iqr)
    tmp <- apply_stable_scaler(val_vec, sc)
    scores_val[[paste0(col, "_stable_z")]] <- tmp$stable_z
    family_scaler_info[[col]] <- sc
  }
  
  for (col in pct_cols) {
    scores_val[[col]] <- val_pos$fam_pct[[col]]
  }
  
  list(
    scores_train = scores_train,
    scores_val = scores_val,
    stable_idx = stable_idx,
    family_axis_cols = fam_cols,
    family_pct_cols = pct_cols,
    family_map = family_map,
    pos_scaler = pos_scaler,
    family_scalers = family_scaler_info
  )
}

# ------------------------------------------------------------
# Dominant family by stable-z and by percentage contribution
# ------------------------------------------------------------
append_family_summaries <- function(scores_val, family_axis_cols, family_pct_cols) {
  z_cols <- paste0(family_axis_cols, "_stable_z")
  
  # dominant family by stable-z
  dom_family <- apply(scores_val[, z_cols, drop = FALSE], 1, function(v) {
    if (all(is.na(v))) return(NA_character_)
    names(v)[which.max(v)]
  })
  
  dom_z <- apply(scores_val[, z_cols, drop = FALSE], 1, function(v) {
    if (all(is.na(v))) return(NA_real_)
    max(v, na.rm = TRUE)
  })
  
  scores_val$dominant_family_axis <- gsub("_axis_stable_z$", "", dom_family)
  scores_val$dominant_family_stable_z <- dom_z
  
  # dominant family by percentage composition
  dom_pct_family <- apply(scores_val[, family_pct_cols, drop = FALSE], 1, function(v) {
    if (all(is.na(v))) return(NA_character_)
    names(v)[which.max(v)]
  })
  
  dom_pct <- apply(scores_val[, family_pct_cols, drop = FALSE], 1, function(v) {
    if (all(is.na(v))) return(NA_real_)
    max(v, na.rm = TRUE)
  })
  
  scores_val$dominant_family_pct_axis <- gsub("_pct$", "", dom_pct_family)
  scores_val$dominant_family_pct <- dom_pct
  
  scores_val
}

# ------------------------------------------------------------
# Qualitative level for overall positive evidence
# ------------------------------------------------------------
positive_evidence_band <- function(z) {
  if (is.na(z)) return("unknown")
  if (z < 0) return("low")
  if (z < 2) return("moderate")
  "high"
}

# ------------------------------------------------------------
# Text summary
# ------------------------------------------------------------
append_risk_profile <- function(scores_val,
                                family_axis_cols,
                                family_pct_cols,
                                top_n_families = 3,
                                d_dist_low_cut = 0,
                                d_dist_high_cut = 5,
                                score_digits = 2,
                                pct_digits = 0) {
  
  missing_cols <- setdiff(c("d_dist", "pos_sum_stable_z", family_pct_cols), names(scores_val))
  if (length(missing_cols) > 0) {
    stop("Missing required columns in scores_val: ", paste(missing_cols, collapse = ", "))
  }
  
  risk_band_from_ddist <- function(x, low_cut = 0, high_cut = 5) {
    if (is.na(x)) return("unknown")
    if (x < low_cut) return("stable / low-risk geometry")
    if (x < high_cut) return("ambiguous / monitor geometry")
    "high-risk geometry"
  }
  
  summarize_top_families_pct <- function(row_vals, pct_names, top_n = 3, digits = 0) {
    vals <- as.numeric(row_vals)
    names(vals) <- gsub("_pct$", "", pct_names)
    
    keep <- is.finite(vals)
    vals <- vals[keep]
    if (length(vals) == 0) return(NA_character_)
    
    ord <- order(vals, decreasing = TRUE)
    ord <- ord[seq_len(min(top_n, length(ord)))]
    
    paste(
      sprintf("%s (%s%%)",
              names(vals)[ord],
              formatC(100 * vals[ord], format = "f", digits = digits)),
      collapse = "; "
    )
  }
  
  profiles <- vapply(seq_len(nrow(scores_val)), function(i) {
    dd <- scores_val$d_dist[i]
    geom_band <- risk_band_from_ddist(dd, d_dist_low_cut, d_dist_high_cut)
    
    pos_z <- scores_val$pos_sum_stable_z[i]
    pos_band <- positive_evidence_band(pos_z)
    
    fam_vals <- scores_val[i, family_pct_cols, drop = FALSE]
    top_fam <- summarize_top_families_pct(
      row_vals = fam_vals,
      pct_names = family_pct_cols,
      top_n = top_n_families,
      digits = pct_digits
    )
    
    paste0(
      "Geometry: ", geom_band,
      " (d_dist=", formatC(dd, format = "f", digits = score_digits), "). ",
      "Overall positive evidence: ", pos_band,
      " (z=", ifelse(is.na(pos_z), "NA", formatC(pos_z, format = "f", digits = score_digits)), "). ",
      "Positive evidence composition: ", top_fam, "."
    )
  }, character(1))
  
  scores_val$geometry_band <- vapply(
    scores_val$d_dist,
    risk_band_from_ddist,
    character(1),
    low_cut = d_dist_low_cut,
    high_cut = d_dist_high_cut
  )
  
  scores_val$positive_evidence_band <- vapply(
    scores_val$pos_sum_stable_z,
    positive_evidence_band,
    character(1)
  )
  
  scores_val$risk_profile <- profiles
  
  scores_val
}

```





```{r}

interp_out <- append_interpretability_scores(
  scores_train = scores_train,
  scores_val = scores_val,
  ll_train = Ltrain,
  ll_val = Lval,
  stable_idx = scores_train$d_dist < 0,
  normalize_family = FALSE,
  normalize_overall = TRUE,
  min_iqr = 1e-4
)

scores_train <- interp_out$scores_train
scores_val   <- interp_out$scores_val

scores_val <- append_family_summaries(
  scores_val = scores_val,
  family_axis_cols = interp_out$family_axis_cols,
  family_pct_cols = interp_out$family_pct_cols
)

scores_val <- append_risk_profile(
  scores_val = scores_val,
  family_axis_cols = interp_out$family_axis_cols,
  family_pct_cols = interp_out$family_pct_cols,
  top_n_families = 3,
  d_dist_low_cut = 0,
  d_dist_high_cut = 2,
  score_digits = 2,
  pct_digits = 0
)

```


### Boxplot of z versus binned d_dist by class. Median separation in z exists between classes in almost every d_dist bin.  

```{r}

scores_val %>% filter(d_dist >= 0, d_dist < 20) %>% mutate(d_dist_bin = cut(d_dist, breaks = seq(0, 20, by = 1))) %>% ggplot() +
  geom_boxplot(aes(y=pos_sum_stable_z, x=d_dist_bin, color=mortality)) +
  theme_minimal()

scores_val %>% filter(d_dist >= 0, d_dist < 20) %>% mutate(d_dist_bin = cut(d_dist, breaks = seq(0, 20, by = 1))) %>% ggplot() +
  geom_boxplot(aes(y=l, x=d_dist_bin, color=mortality)) +
  theme_minimal()

```


```{r}

scores_val %>% filter(d_dist >= 0, d_dist < 20) %>% mutate(pos_sum_bin = cut(pos_sum_stable_z, breaks = seq(0, 20, by = 1))) %>% ggplot() +
  geom_boxplot(aes(y=d_dist, x=pos_sum_bin, color=mortality)) +
  theme_minimal()


```

### Mortality rate and case counts for bins of z  

```{r}

scores_val %>%
  mutate(
    pos_band = cut(
      pos_sum_stable_z,
      breaks = c(-Inf, 0, 1, 2, 3, 4, 5, 6, 7, 8, Inf),
      right = FALSE,
      labels = c("<0", "[0,1)", "[1,2)", "(2,3]","(3,4]", "(4,5]","(5,6]","(6,7]","(7,8]", ">=8")
    )
  ) %>%
  group_by(pos_band) %>%
  summarise(
    count = n(),
    mortality_rate = mean(mortality == 1),
    .groups = "drop"
  )

scores_val %>%
  mutate(
    pos_band = cut(
      pos_sum_stable_z,
      breaks = c(-Inf, 0, 1, 2, 3, 4, 5, 6, 7, 8, Inf),
      right = FALSE,
      labels = c("<0", "[0,1)", "[1,2)", "(2,3]","(3,4]", "(4,5]","(5,6]","(6,7]","(7,8]", ">=8")
    )
  ) %>%
  group_by(pos_band) %>%
  summarise(
    count = n(),
    mortality_rate = mean(mortality == 1),
    .groups = "drop"
  ) %>% ggplot(aes(x=pos_band, y=mortality_rate, group=1)) + geom_point() + geom_line() + theme_minimal()

```


```{r}

pROC::roc(scores_val$mortality, scores_val$pos_sum_stable_z)

```


```{r}

cor(scores_val$p, scores_val$pos_sum_stable_z)

```


```{r}

scores_val%>%filter(d_dist >=0, d_dist < 5) %>% ggplot(aes(x=d_dist, y=l, color=mortality)) + geom_point(alpha=0.6) + theme_minimal()

scores_val%>%filter(d_dist >=0, d_dist < 5) %>% ggplot(aes(x=d_dist, y=pos_sum_stable_z, color=mortality)) + geom_point(alpha=0.6) + theme_minimal()

```


```{r}

pos_sum_val_fit <- glm(pos_sum_stable_z ~ d_dist, data=scores_val)

```

```{r}

summary(pos_sum_val_fit)

```


```{r}

scores_val$pos_sum_cond <- predict(pos_sum_val_fit, scores_val)

```

```{r}

scores_val$risk_axis_residual <- scores_val$pos_sum_stable_z - scores_val$pos_sum_cond

```


```{r}

scores_val%>%filter(d_dist >=0, d_dist < 5) %>% ggplot(aes(x=d_dist, y=risk_axis_residual, color=mortality)) + geom_point(alpha=0.6) + theme_minimal()

```

```{r}

scores_val %>% filter(d_dist >= 0, d_dist < 20) %>% mutate(d_dist_bin = cut(d_dist, breaks = seq(0, 20, by = 1))) %>% ggplot() +
  geom_boxplot(aes(y=risk_axis_residual, x=d_dist_bin, color=mortality)) +
  theme_minimal()

```


## Risk Decomposition

```{r}


mort_0_example <- head(as.vector(head(scores_val%>%filter(d_dist >= 0 & d_dist < 2, mortality=="0"), 1))[grep("_(raw|axis)$", names(as.vector(head(scores_val%>%filter(d_dist >= 0 & d_dist < 2, mortality=="0"), 1))))], -2)

mort_0_example <- mort_0_example[mort_0_example > 0]

mort_0_example <- mort_0_example[order(unlist(mort_0_example), decreasing = TRUE)]

mort_0_example <- head(mort_0_example, 4)

mort_0_example_caption <- head(scores_val%>%filter(d_dist >= 0 & d_dist < 2, mortality=="0"), 1)["risk_profile"]

barplot(as.numeric(mort_0_example), names.arg=names(mort_0_example))
title(sub = "Geometry: ambiguous / monitor geometry (d_dist=0.58).\n Overall positive evidence: moderate (z=0.44).\n Positive evidence composition: spo2 (36%); mbp (18%); sbp (18%)")

```


```{r}


mort_1_example <- head(as.vector(head(scores_val%>%filter(d_dist >= 0 & d_dist < 2, mortality=="1"), 1))[grep("_(raw|axis)$", names(as.vector(head(scores_val%>%filter(d_dist >= 0 & d_dist < 2, mortality=="1"), 1))))], -2)

mort_1_example <- mort_1_example[mort_1_example > 0]

mort_1_example <- mort_1_example[order(unlist(mort_1_example), decreasing = TRUE)]

mort_1_example <- head(mort_1_example, 4)

mort_1_example_caption <- head(scores_val%>%filter(d_dist >= 0 & d_dist < 2, mortality=="1"), 1)["risk_profile"]

barplot(as.numeric(mort_1_example), names.arg=names(mort_1_example))
title(sub = "Geometry: ambiguous / monitor geometry (d_dist=1.60).\n Overall positive evidence: high (z=3.35).\n Positive evidence composition: resp_rate (27%); heart_rate (20%); hemoglobin (11%).")

```


## Evidence Geometry - Test Set

### Apply trained evidence geometry on Test Set

```{r}

Ltest <- loglik_matrices(mimic_clean_test%>%dplyr::select(-stay_id), obj$fit, alpha=spec$alpha)
# val/test scoring
scores_test <- bind_cols(stay_id=mimic_clean_test$stay_id, 
                            mortality=mimic_clean_test$mortality, 
                            score_risk(Ltest$l_pos, Ltest$l_neg,  obj$fit$weights, obj$geom, 
                                       spec$alpha, spec$eps))


```


```{r}

scores_test <- bind_cols(scores_test,
                            p = predict(rf_train_1, mimic_clean_test%>%dplyr::select(-stay_id, -mortality))$predictions[,"1"])

```


### Obtain z and percent contributions of feature groups on Test Set. 

This part uses marginal median and IQR of d_dist < 0 region from training set.  

```{r}

interp_out_test <- append_interpretability_scores(
  scores_train = scores_train,
  scores_val = scores_test,
  ll_train = Ltrain,
  ll_val = Ltest,
  stable_idx = scores_train$d_dist < 0,
  normalize_family = FALSE,
  normalize_overall = TRUE,
  min_iqr = 1e-4
)

scores_train <- interp_out_test$scores_train
scores_test   <- interp_out_test$scores_val

scores_test <- append_family_summaries(
  scores_val = scores_test,
  family_axis_cols = interp_out_test$family_axis_cols,
  family_pct_cols = interp_out_test$family_pct_cols
)

scores_test <- append_risk_profile(
  scores_val = scores_test,
  family_axis_cols = interp_out_test$family_axis_cols,
  family_pct_cols = interp_out_test$family_pct_cols,
  top_n_families = 3,
  d_dist_low_cut = 0,
  d_dist_high_cut = 2,
  score_digits = 2,
  pct_digits = 0
)

```


### Mortality Rate of d_dist < 0 region  

```{r}

nrow(scores_test%>%filter(d_dist < 0 , mortality=="1")) / nrow(scores_test%>%filter(d_dist < 0))

```

### Survived cases in d_dist < 0 as a fraction of total survived cases

```{r}

nrow(scores_test%>%filter(d_dist < 0, mortality=="0")) / nrow(scores_test%>%filter(mortality=="0"))

```

### Mortality Rate curve and case counts  

```{r}

scores_test %>% filter(d_dist >= 0, d_dist < 20) %>% mutate(d_dist_bin = cut(d_dist, breaks = seq(0, 20, by = 1))) %>% group_by(d_dist_bin) %>% dplyr::summarize(count = n(), mortality_rate = mean(mortality == "1"))


scores_test %>% filter(d_dist >= 0, d_dist < 20) %>% mutate(d_dist_bin = cut(d_dist, breaks = seq(0, 20, by = 1))) %>%
  group_by(d_dist_bin) %>%
  summarise(
    mortality_rate = mean(mortality=="1")
  ) %>% ggplot(aes(x=d_dist_bin, y=mortality_rate, group = 1)) + geom_point() + geom_line() + theme_minimal()


```


### Mortality rate and case counts for bins of z  

```{r}

scores_test %>%
  mutate(
    pos_band = cut(
      pos_sum_stable_z,
      breaks = c(-Inf, 0, 1, 2, 3, 4, 5, 6, 7, 8, Inf),
      right = FALSE,
      labels = c("<0", "[0,1)", "[1,2)", "(2,3]","(3,4]", "(4,5]","(5,6]","(6,7]","(7,8]", ">=8")
    )
  ) %>%
  group_by(pos_band) %>%
  summarise(
    count = n(),
    mortality_rate = mean(mortality == 1),
    .groups = "drop"
  )

scores_test %>%
  mutate(
    pos_band = cut(
      pos_sum_stable_z,
      breaks = c(-Inf, 0, 1, 2, 3, 4, 5, 6, 7, 8, Inf),
      right = FALSE,
      labels = c("<0", "[0,1)", "[1,2)", "(2,3]","(3,4]", "(4,5]","(5,6]","(6,7]","(7,8]", ">=8")
    )
  ) %>%
  group_by(pos_band) %>%
  summarise(
    count = n(),
    mortality_rate = mean(mortality == 1),
    .groups = "drop"
  ) %>% ggplot(aes(x=pos_band, y=mortality_rate, group=1)) + geom_point() + geom_line() + theme_minimal()

```

### Triage Table
  
1. Boxplot of z versus d_dist bin by class.  
2. Primary triage table using d_dist ($d_{dist} \geq 2$ -> High Risk)  
  
```{r}

scores_test %>% filter(d_dist >= 0, d_dist < 20) %>% mutate(d_dist_bin = cut(d_dist, breaks = seq(0, 20, by = 1))) %>% ggplot() +
  geom_boxplot(aes(y=pos_sum_stable_z, x=d_dist_bin, color=mortality)) +
  theme_minimal()

scores_test %>% group_by(geometry_band) %>% dplyr::summarize(count = n(), mortality_rate = mean(mortality=="1"), percent_pop = n()/nrow(scores_test))

```


### Example Risk Profiles

```{r}

head(scores_test$risk_profile, 8)

```

### RF p vs mortality rate

```{r}

scores_test %>% mutate(p_bin = cut(p, breaks = 25)) %>% group_by(p_bin) %>% dplyr::summarize(count = n(), mortality_rate = mean(mortality == "1"))


scores_test %>% mutate(p_bin = cut(p, breaks = 25)) %>%
  group_by(p_bin) %>%
  summarise(
    mortality_rate = mean(mortality=="1")
  ) %>% ggplot(aes(x=p_bin, y=mortality_rate, group = 1)) + geom_point() + geom_line() + theme_minimal()


```

### RF mortality in matched low risk zone size (17.13%)

```{r}

rf_matched_safe_zone <- head(scores_test%>%select(mortality, p)%>%arrange(p), round(nrow(scores_test)*0.1713065,0))

```

```{r}

nrow(rf_matched_safe_zone%>%filter(mortality=="1"))/nrow(rf_matched_safe_zone)

```

```{r}

max(rf_matched_safe_zone$p)

```


### RF mortality in matched high risk zone size (36.65%)

```{r}

rf_matched_highrisk_zone <- tail(scores_test%>%select(mortality, p)%>%arrange(p), round(0.3664862*nrow(scores_test),0))

```

```{r}

nrow(rf_matched_highrisk_zone%>%filter(mortality=="1"))/nrow(rf_matched_highrisk_zone)

```


## Survival Analysis

```{r}

mimiciv_cohort <- read.csv(here("v2","src","data","mimiciv","mimiciv_ext_icu_cohort_1_timetodeath.csv"))

```

```{r}

mimiciv_timeperiod <- mimiciv_cohort%>%dplyr::select(stay_id, duration_hours_from_24h, event_after_24h)

```

```{r}

cohort_age <- mimiciv_import%>%dplyr::select(stay_id, anchor_age)

```

```{r}

mimiciv_timeperiod$duration_hours_from_24h <- ifelse(
  mimiciv_timeperiod$duration_hours_from_24h <= 0,
  0.01,
  mimiciv_timeperiod$duration_hours_from_24h
)

```


```{r}

risk_timeperiod_train <- left_join(scores_train, mimiciv_timeperiod, by=join_by(stay_id))%>%dplyr::select(all_of(c(names(scores_train), "duration_hours_from_24h", "event_after_24h")))

risk_timeperiod_train <- left_join(risk_timeperiod_train, cohort_age, by=join_by(stay_id))


```

```{r}

risk_timeperiod_val <- left_join(scores_val, mimiciv_timeperiod, by=join_by(stay_id))%>%dplyr::select(all_of(c(names(scores_val), "duration_hours_from_24h", "event_after_24h")))

risk_timeperiod_val <- left_join(risk_timeperiod_val, cohort_age, by=join_by(stay_id))

```


```{r}

risk_timeperiod_test <- left_join(scores_test, mimiciv_timeperiod, by=join_by(stay_id))%>%dplyr::select(all_of(c(names(scores_test), "duration_hours_from_24h", "event_after_24h")))

risk_timeperiod_test <- left_join(risk_timeperiod_test, cohort_age, by=join_by(stay_id))

```



### Kaplan-Meier Analysis of d_dist


```{r}

risk_timeperiod_train$d_dist_bin <- cut(
  risk_timeperiod_train$d_dist,
  breaks = c(-Inf, -0.1, 2, Inf),
  labels = c("Low", "Ambiguous", "High")
)

```


```{r}


km_fit <- survfit(
  Surv(duration_hours_from_24h, event_after_24h) ~ d_dist_bin,
  data = risk_timeperiod_train
)

ggsurvplot(
  km_fit,
  data = risk_timeperiod_train,
  risk.table = TRUE,
  pval = TRUE
)

```

### Age Only

```{r}

M_base <- coxph(Surv(duration_hours_from_24h, event_after_24h) ~ anchor_age, data = risk_timeperiod_train)

summary(M_base)

```



```{r}

lp_age <- predict(M_base, newdata = risk_timeperiod_val, type = "lp")

# Concordance
concordance(Surv(duration_hours_from_24h, event_after_24h) ~ lp_age, 
            data = risk_timeperiod_val, reverse = TRUE)

# timeROC
time_roc_age <- timeROC(
  T = risk_timeperiod_val$duration_hours_from_24h,
  delta = risk_timeperiod_val$event_after_24h,
  marker = lp_age,
  cause = 1,
  times = c(48, 60, 72, 96, 168),
  iid = FALSE
)

time_roc_age$AUC

```

### d_dist

```{r}

M_geom <- coxph(Surv(duration_hours_from_24h, event_after_24h) ~ d_dist, data = risk_timeperiod_train)

summary(M_geom)

```

```{r}

lp_geom <- predict(M_geom, newdata = risk_timeperiod_val, type = "lp")

# Concordance
concordance(Surv(duration_hours_from_24h, event_after_24h) ~ lp_geom, 
            data = risk_timeperiod_val, reverse = TRUE)

# timeROC
time_roc_geom <- timeROC(
  T = risk_timeperiod_val$duration_hours_from_24h,
  delta = risk_timeperiod_val$event_after_24h,
  marker = lp_geom,
  cause = 1,
  times = c(48, 60, 72, 96, 168),
  iid = FALSE
)

time_roc_geom$AUC

```

```{r}

M_severity <- coxph(Surv(duration_hours_from_24h, event_after_24h) ~ pos_sum, data = risk_timeperiod_train)

summary(M_severity)

```

```{r}

M_geom_severity <- coxph(Surv(duration_hours_from_24h, event_after_24h) ~ d_dist + pos_sum, data = risk_timeperiod_train)

summary(M_geom_severity) 

```


```{r}

anova(M_base, M_geom, M_severity, M_geom_severity, test="LRT")

```

### Measurement Groups Only

```{r}

M_composition <- coxph(as.formula(paste("Surv(duration_hours_from_24h, event_after_24h) ~ ", 
                             paste(names(scores_train)[grep("_axis", names(scores_train))], collapse = " + "), collapse='')), 
                       data = risk_timeperiod_train)

summary(M_composition)

```


```{r}

lp_comp <- predict(M_composition, newdata = risk_timeperiod_val, type = "lp")

# Concordance
concordance(Surv(duration_hours_from_24h, event_after_24h) ~ lp_comp, 
            data = risk_timeperiod_val, reverse = TRUE)

# timeROC
time_roc_comp <- timeROC(
  T = risk_timeperiod_val$duration_hours_from_24h,
  delta = risk_timeperiod_val$event_after_24h,
  marker = lp_comp,
  cause = 1,
  times = c(48, 60, 72, 96, 168),
  iid = FALSE
)

time_roc_comp$AUC

```

```{r}

anova(M_base, M_geom_severity, M_composition, test="LRT")

```


```{r}

M_d_dist_composition <- coxph(as.formula(paste("Surv(duration_hours_from_24h, event_after_24h) ~ d_dist +", 
                             paste(names(scores_train)[grep("_axis", names(scores_train))], collapse = " + "), collapse='')), 
                       data = risk_timeperiod_train)

summary(M_d_dist_composition)

```

```{r}

M_d_dist_pos_sum_composition <- coxph(as.formula(paste("Surv(duration_hours_from_24h, event_after_24h) ~ d_dist + pos_sum + ", 
                             paste(names(scores_train)[grep("_axis", names(scores_train))], collapse = " + "), collapse='')), 
                       data = risk_timeperiod_train)

summary(M_d_dist_pos_sum_composition)

```


```{r}

anova(M_base, M_composition, M_d_dist_composition, test="LRT")

```

```{r}

M_d_dist_strata_composition <- coxph(as.formula(paste("Surv(duration_hours_from_24h, event_after_24h) ~ ", 
                             paste(names(scores_train)[grep("_axis", names(scores_train))], collapse = " + "), 
                             " + strata(d_dist_bin)", collapse='')), 
                       data = risk_timeperiod_train)

summary(M_d_dist_strata_composition)

```


### d_dist + axes

```{r}

lp_train_raw <- predict(M_d_dist_composition, newdata = risk_timeperiod_train, type = "lp", reference = "zero")
lp_val_raw   <- predict(M_d_dist_composition, newdata = risk_timeperiod_val,   type = "lp", reference = "zero")

```

```{r}

lp_anchor <- mean(lp_train_raw[risk_timeperiod_train$d_dist < 0], na.rm = TRUE)

```

```{r}

risk_timeperiod_train$lp_rel <- lp_train_raw - lp_anchor
risk_timeperiod_val$lp_rel   <- lp_val_raw   - lp_anchor

risk_timeperiod_train$rr_rel <- exp(risk_timeperiod_train$lp_rel)
risk_timeperiod_val$rr_rel   <- exp(risk_timeperiod_val$lp_rel)

```

```{r}

concordance(Surv(duration_hours_from_24h, event_after_24h) ~ lp_rel, data = risk_timeperiod_val, reverse = TRUE)

```

```{r}

time_roc <- timeROC(
  T = risk_timeperiod_val$duration_hours_from_24h,
  delta = risk_timeperiod_val$event_after_24h,
  marker = risk_timeperiod_val$lp_rel,   # or rr_rel, same ranking
  cause = 1,
  times = c(48, 60, 72, 96, 168),
  iid = FALSE
)

time_roc$AUC

```

```{r}

risk_timeperiod_val$risk_group <- cut(risk_timeperiod_val$lp_rel,
                      breaks = quantile(risk_timeperiod_val$lp_rel, probs = c(0, 0.25, 0.5, 0.75, 1)),
                      include.lowest = TRUE)

risk_fit_val <- survfit(Surv(duration_hours_from_24h, event_after_24h) ~ risk_group, data = risk_timeperiod_val)
plot(risk_fit_val, col = 1:4, lwd = 2)

```


```{r}

termplot(M_d_dist_composition, se = TRUE)

```


```{r}

dd <- datadist(risk_timeperiod_train)
options(datadist = "dd")

```


```{r}

cox_spline1 <- cph(
  Surv(duration_hours_from_24h, event_after_24h) ~ 
    rcs(d_dist, 4) +
    anchor_age_axis +
    gender_axis +
    bicarbonate_axis +
    resp_rate_axis +
    lactate_axis +
    creatinine_axis +
    platelet_axis +
    spo2_axis + 
    sbp_axis +
    temperature_axis,
  data = risk_timeperiod_train,
  x = TRUE,
  y = TRUE,
  surv = TRUE
)

```


```{r}

cox_spline2 <- cph(
  Surv(duration_hours_from_24h, event_after_24h) ~ 
    rcs(d_dist, 4),
  data = risk_timeperiod_train,
  x = TRUE,
  y = TRUE,
  surv = TRUE
)

```


```{r}

plot(Predict(cox_spline1, d_dist), 
     xlab = "d_dist", 
     ylab = "Log-Hazard Ratio")

plot(Predict(cox_spline1, anchor_age_axis), 
     xlab = "anchor_age_axis", 
     ylab = "Log-Hazard Ratio")

plot(Predict(cox_spline1, resp_rate_axis), 
     xlab = "resp_rate_axis", 
     ylab = "Log-Hazard Ratio")

```

```{r}

plot(Predict(cox_spline2, d_dist), 
     xlab = "d_dist", 
     ylab = "Log-Hazard Ratio")


```

### Spline(d_dist) + axes

```{r}

lp_spline_train_raw <- predict(cox_spline1, newdata = risk_timeperiod_train, type = "lp")
lp_spline_val_raw   <- predict(cox_spline1, newdata = risk_timeperiod_val,   type = "lp")

```


```{r}

lp_spline_anchor <- mean(lp_spline_train_raw[risk_timeperiod_train$d_dist < 0], na.rm = TRUE)

risk_timeperiod_val$lp_spline_rel <- lp_spline_val_raw - lp_spline_anchor

```


```{r}

# Concordance
concordance(Surv(duration_hours_from_24h, event_after_24h) ~ risk_timeperiod_val$lp_spline_rel, 
            data = risk_timeperiod_val, reverse = TRUE)

# timeROC
time_roc_spline <- timeROC(
  T = risk_timeperiod_val$duration_hours_from_24h,
  delta = risk_timeperiod_val$event_after_24h,
  marker = risk_timeperiod_val$lp_spline_rel,
  cause = 1,
  times = c(48, 60, 72, 96, 168),
  iid = FALSE
)

time_roc_spline$AUC

```

### Spline(d_dist)

```{r}

lp_spline2_train_raw <- predict(cox_spline2, newdata = risk_timeperiod_train, type = "lp")
lp_spline2_val_raw   <- predict(cox_spline2, newdata = risk_timeperiod_val,   type = "lp")

```


```{r}

lp_spline2_anchor <- mean(lp_spline2_train_raw[risk_timeperiod_train$d_dist < 0], na.rm = TRUE)

risk_timeperiod_val$lp_spline2_rel <- lp_spline2_val_raw - lp_spline2_anchor

```

```{r}

# Concordance
concordance(Surv(duration_hours_from_24h, event_after_24h) ~ risk_timeperiod_val$lp_spline2_rel, 
            data = risk_timeperiod_val, reverse = TRUE)

# timeROC
time_roc_spline2 <- timeROC(
  T = risk_timeperiod_val$duration_hours_from_24h,
  delta = risk_timeperiod_val$event_after_24h,
  marker = risk_timeperiod_val$lp_spline2_rel,
  cause = 1,
  times = c(48, 60, 72, 96, 168),
  iid = FALSE
)

time_roc_spline2$AUC

```


```{r}

lp_spline_terms <- predict(cox_spline1, newdata = risk_timeperiod_val, type = "terms")

```

### Raw Feature Subset

```{r}

cox_baseline_train <- left_join(mimic_clean_train, mimiciv_timeperiod, by=join_by(stay_id))

```

```{r}

cox_baseline_val <- left_join(mimic_clean_val, mimiciv_timeperiod, by=join_by(stay_id))

```


```{r}

raw_vars <- c(
  "anchor_age",
  "gender",
  "heart_rate_mean",
  "resp_rate_mean",
  "sbp_mean",
  "dbp_mean",
  "mbp_mean",
  "spo2_mean",
  "temperature_mean",
  "glucose_max",
  "creatinine_max",
  "bicarbonate_min",
  "lactate_max",
  "sodium_min",
  "potassium_max",
  "platelet_min",
  "wbc_max",
  "hemoglobin_min"
)

form_raw <- as.formula(
  paste(
    "Surv(duration_hours_from_24h, event_after_24h) ~",
    paste(raw_vars, collapse = " + ")
  )
)

cox_raw <- coxph(form_raw, data = cox_baseline_train)
summary(cox_raw)

```


```{r}

lp_raw_val <- predict(cox_raw, newdata = cox_baseline_val, type = "lp")

concordance(
  Surv(duration_hours_from_24h, event_after_24h) ~ lp_raw_val,
  data = cox_baseline_val,
  reverse = TRUE
)

time_roc_baseline <- timeROC(
  T = risk_timeperiod_val$duration_hours_from_24h,
  delta = risk_timeperiod_val$event_after_24h,
  marker = lp_raw_val,
  cause = 1,
  times = c(48, 72, 96, 168),
  iid = FALSE
)

time_roc_baseline$AUC

```


## Age Ablation

Analysis to check influence of age in survival model

```{r}

raw_vars2 <- c(
  "gender",
  "heart_rate_mean",
  "resp_rate_mean",
  "sbp_mean",
  "dbp_mean",
  "mbp_mean",
  "spo2_mean",
  "temperature_mean",
  "glucose_max",
  "creatinine_max",
  "bicarbonate_min",
  "lactate_max",
  "sodium_min",
  "potassium_max",
  "platelet_min",
  "wbc_max",
  "hemoglobin_min"
)

form_raw2 <- as.formula(
  paste(
    "Surv(duration_hours_from_24h, event_after_24h) ~",
    paste(raw_vars2, collapse = " + ")
  )
)

cox_raw2 <- coxph(form_raw2, data = cox_baseline_train)
summary(cox_raw2)

```

```{r}

lp_raw_val2 <- predict(cox_raw2, newdata = cox_baseline_val, type = "lp")

concordance(
  Surv(duration_hours_from_24h, event_after_24h) ~ lp_raw_val2,
  data = cox_baseline_val,
  reverse = TRUE
)

time_roc_baseline2 <- timeROC(
  T = risk_timeperiod_val$duration_hours_from_24h,
  delta = risk_timeperiod_val$event_after_24h,
  marker = lp_raw_val2,
  cause = 1,
  times = c(48, 72, 96, 168),
  iid = FALSE
)

time_roc_baseline2$AUC

```


```{r}

mimic_clean_train2 <- mimic_clean_train%>%select(-anchor_age)

mimic_clean_val2 <- mimic_clean_val%>%select(-anchor_age)

```


```{r}

spec2 <- risk_spec(
   y_col="mortality", positive="1",
   features = setdiff(names(mimic_clean_train2), c("mortality", "stay_id")),
   alpha = alpha,
   laplace = 1, ridge = 1e-3, winsor_p=0.01,
   weights=FALSE,
   weight_method = "mutual_info",
   numeric_candidates = c("gaussian", "lognormal", "gamma"),
   count_candidates = c("poisson", "negbinom"),
   fraction_candidates = c("logit_gaussian"),
   fraction_eps = 1e-3,
   numeric_val_frac = 0.2,
   numeric_min_n = 25,
   llr_cap_quantile = 0.01,
   mi_nbins = 10
 )

obj2 <- fit(mimic_clean_train2%>%dplyr::select(-stay_id), spec2, k_eigen=4, k_energy=4, energy_ref="both")


```

### Compute Log-Likelihood Ratio (LLR) Matrix for Validation Set. Compute geometric risk signals

```{r}

Lval2 <- loglik_matrices(mimic_clean_val2%>%dplyr::select(-stay_id), obj2$fit, alpha=spec2$alpha)
# val/test scoring
scores_val2 <- bind_cols(stay_id=mimic_clean_val2$stay_id, 
                            mortality=mimic_clean_val2$mortality, 
                            score_risk(Lval2$l_pos, Lval2$l_neg,  obj2$fit$weights, obj2$geom, 
                                       spec2$alpha, spec2$eps))


```

### Train Random Forest on Train Set

```{r}

rf_train_2 <- ranger(formula = mortality ~ ., data=mimic_clean_train2%>%dplyr::select(-stay_id),
                         mtry=10, min.node.size=7, num.trees=500, probability = TRUE,
                         keep.inbag = TRUE)

```

### Append RF predictions on Validation set to geometric risk dataframe

```{r}

scores_val2 <- bind_cols(scores_val2,
                            p = predict(rf_train_2, mimic_clean_val2%>%dplyr::select(-stay_id, -mortality))$predictions[,"1"])

```



```{r}

ggplot(scores_val2) +
  geom_density(aes(x=p, fill=mortality), alpha=0.6) +
  xlab("Random Forest probability of mortality") +
  theme_minimal()

ggplot(scores_val2) +
  geom_density(aes(x=l, fill=mortality), alpha=0.6) +
  xlab("Sum of relative marginal evidence from all features") +
  theme_minimal()

ggplot(scores_val2) +
  geom_density(aes(x=d_dist, fill=mortality), alpha=0.6) +
  xlab("d_dist") +
  theme_minimal()

ggplot(scores_val2) +
  geom_density(aes(x=proj, fill=mortality), alpha=0.6) +
  xlab("proj") +
  theme_minimal()

ggplot(scores_val2) +
  geom_point(aes(x=l, y=p, color=mortality), alpha=0.4) +
  ylab("Random forest mortality probability (p)") +
  theme_minimal()

ggplot(scores_val2) +
  geom_point(aes(x=qlogis(p), y=l, color=mortality), alpha=0.4) +
  xlab("log-odds of random forest mortality probability (p)") +
  theme_minimal()

ggplot(scores_val2) +
  geom_point(aes(x=p, y=d_dist, color=mortality), alpha=0.4) +
  xlab("Random forest mortality probability (p)") +
  theme_minimal()

ggplot(scores_val2) +
  geom_point(aes(x=p, y=proj, color=mortality), alpha=0.4) +
  xlab("Random forest mortality probability (p)") +
  theme_minimal()

ggplot(scores_val2) +
  geom_point(aes(x=d_dist, y=proj, color=mortality), alpha=0.4) +
  theme_minimal()

ggplot(scores_val2%>%filter(d_dist > -4, d_dist < 4)) +
  geom_point(aes(x=d_dist, y=proj, color=mortality), alpha=0.4) +
  theme_minimal()


```


```{r}

Ltrain2 <- loglik_matrices(mimic_clean_train2%>%dplyr::select(-stay_id), obj2$fit, alpha=spec2$alpha)

scores_train2 <- bind_cols(stay_id=mimic_clean_train2$stay_id, 
                            mortality=mimic_clean_train2$mortality, 
                            score_risk(Ltrain2$l_pos, Ltrain2$l_neg,  obj2$fit$weights, obj2$geom, 
                                       spec2$alpha, spec2$eps))

```

```{r}

interp_out2 <- append_interpretability_scores(
  scores_train = scores_train2,
  scores_val = scores_val2,
  ll_train = Ltrain2,
  ll_val = Lval2,
  stable_idx = scores_train2$d_dist < 0,
  normalize_family = TRUE,
  normalize_overall = TRUE,
  min_iqr = 1e-4
)

scores_train2 <- interp_out2$scores_train
scores_val2   <- interp_out2$scores_val

scores_val2 <- append_family_summaries(
  scores_val = scores_val2,
  family_axis_cols = interp_out2$family_axis_cols,
  family_pct_cols = interp_out2$family_pct_cols
)

scores_val2 <- append_risk_profile(
  scores_val = scores_val2,
  family_axis_cols = interp_out2$family_axis_cols,
  family_pct_cols = interp_out2$family_pct_cols,
  top_n_families = 3,
  d_dist_low_cut = 0,
  d_dist_high_cut = 2,
  score_digits = 2,
  pct_digits = 0
)

```


```{r}



```

```{r}

risk_timeperiod_train2 <- left_join(scores_train2, mimiciv_timeperiod, by=join_by(stay_id))%>%dplyr::select(all_of(c(names(scores_train2), "duration_hours_from_24h", "event_after_24h")))

risk_timeperiod_val2 <- left_join(scores_val2, mimiciv_timeperiod, by=join_by(stay_id))%>%dplyr::select(all_of(c(names(scores_val2), "duration_hours_from_24h", "event_after_24h")))

```

```{r}

cox_spline1_no_age <- cph(
  Surv(duration_hours_from_24h, event_after_24h) ~ 
    rcs(d_dist, 4) +
    gender_axis +
    bicarbonate_axis +
    resp_rate_axis +
    lactate_axis +
    creatinine_axis +
    platelet_axis +
    spo2_axis + 
    sbp_axis +
    temperature_axis,
  data = risk_timeperiod_train2,
  x = TRUE,
  y = TRUE,
  surv = TRUE
)

```




```{r}

lp_spline_no_age_train_raw <- predict(cox_spline1_no_age, newdata = risk_timeperiod_train2, type = "lp")
lp_spline_no_age_val_raw   <- predict(cox_spline1_no_age, newdata = risk_timeperiod_val2,   type = "lp")

```


```{r}

lp_spline_no_age_anchor <- mean(lp_spline_no_age_train_raw[risk_timeperiod_train2$d_dist < 0], na.rm = TRUE)

risk_timeperiod_val2$lp_spline_no_age_rel <- lp_spline_no_age_val_raw - lp_spline_no_age_anchor

```


```{r}

# Concordance
concordance(Surv(duration_hours_from_24h, event_after_24h) ~ risk_timeperiod_val2$lp_spline_no_age_rel, 
            data = risk_timeperiod_val2, reverse = TRUE)

# timeROC
time_roc_spline_no_age <- timeROC(
  T = risk_timeperiod_val2$duration_hours_from_24h,
  delta = risk_timeperiod_val2$event_after_24h,
  marker = risk_timeperiod_val2$lp_spline_no_age_rel,
  cause = 1,
  times = c(48, 60, 72, 96, 168),
  iid = FALSE
)

time_roc_spline_no_age$AUC

```


## ICU Scores Comparison

```{r}

mimiciv_icu_scores <- read.csv(here("v2","src","data","mimiciv","mimiciv_ext_icu_scores.csv"))

```

```{r}

mortality <- mimiciv_import%>%dplyr::select(stay_id, mortality)

```


```{r}

mimiciv_icu_scores <- left_join(mimiciv_icu_scores, cohort_age, by=join_by(stay_id))

mimiciv_icu_scores <- left_join(mimiciv_icu_scores, mimiciv_timeperiod, by=join_by(stay_id))

```


```{r}

mimiciv_icu_scores_train <- inner_join(mimic_clean_train%>%dplyr::select(stay_id), mimiciv_icu_scores, by=join_by(stay_id))

```

```{r}

mimiciv_icu_scores_val <- inner_join(mimic_clean_val%>%dplyr::select(stay_id), mimiciv_icu_scores, by=join_by(stay_id))

```

### SAPSII

```{r}

surv_sapsii <- coxph(Surv(duration_hours_from_24h, event_after_24h) ~ sapsii, data = mimiciv_icu_scores_train)
summary(surv_sapsii)
lp_surv_sapsii <- predict(surv_sapsii, newdata = mimiciv_icu_scores_val, type = "lp")

concordance(
  Surv(duration_hours_from_24h, event_after_24h) ~ lp_surv_sapsii,
  data = mimiciv_icu_scores_val,
  reverse = TRUE
)

time_roc_sapsii <- timeROC(
  T = mimiciv_icu_scores_val$duration_hours_from_24h,
  delta = mimiciv_icu_scores_val$event_after_24h,
  marker = lp_surv_sapsii,
  cause = 1,
  times = c(48, 72, 96, 168),
  iid = FALSE
)

time_roc_sapsii$AUC

```

### APSIII

```{r}

surv_apsiii <- coxph(Surv(duration_hours_from_24h, event_after_24h) ~ anchor_age + apsiii, data = mimiciv_icu_scores_train)
summary(surv_apsiii)
lp_surv_apsiii <- predict(surv_apsiii, newdata = mimiciv_icu_scores_val, type = "lp")

concordance(
  Surv(duration_hours_from_24h, event_after_24h) ~ lp_surv_apsiii,
  data = mimiciv_icu_scores_val,
  reverse = TRUE
)

time_roc_apsiii <- timeROC(
  T = mimiciv_icu_scores_val$duration_hours_from_24h,
  delta = mimiciv_icu_scores_val$event_after_24h,
  marker = lp_surv_apsiii,
  cause = 1,
  times = c(48, 72, 96, 168),
  iid = FALSE
)

time_roc_apsiii$AUC

```

### OASIS

```{r}

surv_oasis <- coxph(Surv(duration_hours_from_24h, event_after_24h) ~ oasis, data = mimiciv_icu_scores_train)
summary(surv_oasis)
lp_surv_oasis <- predict(surv_oasis, newdata = mimiciv_icu_scores_val, type = "lp")

concordance(
  Surv(duration_hours_from_24h, event_after_24h) ~ lp_surv_oasis,
  data = mimiciv_icu_scores_val,
  reverse = TRUE
)

time_roc_oasis <- timeROC(
  T = mimiciv_icu_scores_val$duration_hours_from_24h,
  delta = mimiciv_icu_scores_val$event_after_24h,
  marker = lp_surv_oasis,
  cause = 1,
  times = c(48, 72, 96, 168),
  iid = FALSE
)

time_roc_oasis$AUC

```

### SIRS

```{r}

surv_sirs <- coxph(Surv(duration_hours_from_24h, event_after_24h) ~ anchor_age + sirs, data = mimiciv_icu_scores_train)
summary(surv_sirs)
lp_surv_sirs <- predict(surv_sirs, newdata = mimiciv_icu_scores_val, type = "lp")

concordance(
  Surv(duration_hours_from_24h, event_after_24h) ~ lp_surv_sirs,
  data = mimiciv_icu_scores_val,
  reverse = TRUE
)

time_roc_sirs <- timeROC(
  T = mimiciv_icu_scores_val$duration_hours_from_24h,
  delta = mimiciv_icu_scores_val$event_after_24h,
  marker = lp_surv_sirs,
  cause = 1,
  times = c(48, 72, 96, 168),
  iid = FALSE
)

time_roc_sirs$AUC

```

### LODS

```{r}

surv_lods <- coxph(Surv(duration_hours_from_24h, event_after_24h) ~ anchor_age + lods, data = mimiciv_icu_scores_train)
summary(surv_lods)
lp_surv_lods <- predict(surv_lods, newdata = mimiciv_icu_scores_val, type = "lp")

concordance(
  Surv(duration_hours_from_24h, event_after_24h) ~ lp_surv_lods,
  data = mimiciv_icu_scores_val,
  reverse = TRUE
)

time_roc_lods <- timeROC(
  T = mimiciv_icu_scores_val$duration_hours_from_24h,
  delta = mimiciv_icu_scores_val$event_after_24h,
  marker = lp_surv_lods,
  cause = 1,
  times = c(48, 72, 96, 168),
  iid = FALSE
)

time_roc_lods$AUC

```

### SOFA

```{r}

surv_sofa <- coxph(Surv(duration_hours_from_24h, event_after_24h) ~ anchor_age + sofa_24hours, data = mimiciv_icu_scores_train)
summary(surv_sofa)
lp_surv_sofa <- predict(surv_sofa, newdata = mimiciv_icu_scores_val, type = "lp")

concordance(
  Surv(duration_hours_from_24h, event_after_24h) ~ lp_surv_sofa,
  data = mimiciv_icu_scores_val,
  reverse = TRUE
)

time_roc_sofa <- timeROC(
  T = mimiciv_icu_scores_val$duration_hours_from_24h,
  delta = mimiciv_icu_scores_val$event_after_24h,
  marker = lp_surv_sofa,
  cause = 1,
  times = c(48, 72, 96, 168),
  iid = FALSE
)

time_roc_sofa$AUC

```


## SURVIVAL ANALYSIS TABLE


```{r}

survival_analysis_table <-
  data.frame(
    method = c("Age", "SAPSII", "APSIII + Age", "OASIS", "SIRS + Age", "LODS + Age", "SOFA 24h + Age", "Raw Measurement Subset",
               "d_dist", "d_dist + Measurement Groups", "Spline(d_dist)", "Spline(d_dist) + Measurement Groups"),
    concordance = c(0.6182, 0.7687, 0.7657, 0.7377, 0.6636, 0.7524, 0.7344, 0.7444, 0.7646, 0.7768, 0.7646, 0.7911),
    AUROC_48h = c(0.621, 0.821, 0.815, 0.800, 0.689, 0.788, 0.771, 0.807, 0.836, 0.836, 0.836, 0.857),
    AUROC_72h = c(0.622, 0.802, 0.791, 0.774, 0.672, 0.787, 0.751, 0.788, 0.813, 0.814, 0.813, 0.835),
    AUROC_96h = c(0.616, 0.774, 0.766, 0.752, 0.663, 0.750, 0.730, 0.760, 0.778, 0.788, 0.778, 0.805),
    AUROC_168h = c(0.630, 0.733, 0.737, 0.698, 0.661, 0.722, 0.712, 0.729, 0.713, 0.747, 0.713, 0.754)
  )

survival_analysis_table

```


Raw Feature Subset: c(
  "anchor_age",
  "gender",
  "heart_rate_mean",
  "resp_rate_mean",
  "sbp_mean",
  "dbp_mean",
  "mbp_mean",
  "spo2_mean",
  "temperature_mean",
  "glucose_max",
  "creatinine_max",
  "bicarbonate_min",
  "lactate_max",
  "sodium_min",
  "potassium_max",
  "platelet_min",
  "wbc_max",
  "hemoglobin_min"
)


```{r}

risk_timeperiod_val$risk_group_spline <- cut(risk_timeperiod_val$lp_spline_rel,
                      breaks = quantile(risk_timeperiod_val$lp_spline_rel, probs = c(0, 0.25, 0.5, 0.75, 1)),
                      include.lowest = TRUE)

risk_fit_spline_val <- survfit(Surv(duration_hours_from_24h, event_after_24h) ~ risk_group_spline, data = risk_timeperiod_val)
plot(risk_fit_spline_val, col = 1:4, lwd = 2)

```

```{r}

cox.zph(cox_spline1)

```

```{r}

table(risk_timeperiod_val$risk_group_spline, risk_timeperiod_val$event_after_24h)

```




## Random Forest Performance Enhancement Analysis

Analysis of performance of random forest when trained on evidence data

```{r}

rf_train_3 <- ranger(formula = mortality ~ ., data = bind_cols(mortality=mimic_clean_train$mortality, Ltrain$L),
                     mtry=10, min.node.size=7, num.trees=500, probability = TRUE,
                     keep.inbag = TRUE)

```


```{r}

rf_on_evidence_val <- bind_cols(mortality=mimic_clean_val$mortality, p=predict(rf_train_3, Lval$L)$predictions[,"1"])

```


```{r}

rf_on_evidence_val %>% ggplot(aes(x=p, fill=mortality)) + geom_density(alpha=0.6) + theme_minimal()

```


```{r}

pROC::roc(rf_on_evidence_val$mortality, rf_on_evidence_val$p)

```


## Risk Decomposition Plots - Hazard Ratio


```{r}

terms_mat <- predict(cox_spline1, newdata = risk_timeperiod_val, type = "terms")

```


```{r}

i <- match("0", (scores_val%>%filter(d_dist >= 0 & d_dist < 2))$mortality)
patient_terms <- terms_mat[i, ]
patient_terms

```

```{r}

patient_hr <- exp(patient_terms)
patient_hr

```

```{r}

haz_df <- tibble(
  term = names(patient_terms),
  log_hazard_contrib = as.numeric(patient_terms)
) %>%
  mutate(
    hazard_multiplier = exp(log_hazard_contrib)
  ) %>%
  arrange(log_hazard_contrib)

```


```{r}

ggplot(haz_df, aes(x = reorder(term, hazard_multiplier), y = hazard_multiplier)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Patient-level Hazard Multipliers",
    x = NULL,
    y = "Hazard multiplier"
  )

```

```{r}

j <- match("1", (scores_val%>%filter(d_dist >= 0 & d_dist < 2))$mortality)

patient_terms2 <- terms_mat[j, ]
patient_terms2

haz_df2 <- tibble(
  term = names(patient_terms2),
  log_hazard_contrib = as.numeric(patient_terms2)
) %>%
  mutate(
    hazard_multiplier = exp(log_hazard_contrib)
  ) %>%
  arrange(log_hazard_contrib)

```

```{r}

ggplot(haz_df2, aes(x = reorder(term, hazard_multiplier), y = hazard_multiplier)) +
  geom_col() +
  coord_flip() +
  labs(
    title = "Patient-level Hazard Multipliers",
    x = NULL,
    y = "Hazard multiplier"
  )

```


## References

Johnson, A., Bulgarelli, L., Pollard, T., Gow, B., Moody, B., Horng, S., Celi, L. A., & Mark, R. (2024). MIMIC-IV (version 3.1). PhysioNet. RRID:SCR_007345. https://doi.org/10.13026/kpb9-mt58  
  
Johnson, A.E.W., Bulgarelli, L., Shen, L. et al. MIMIC-IV, a freely accessible electronic health record dataset. Sci Data 10, 1 (2023). https://doi.org/10.1038/s41597-022-01899-x  
  
Goldberger, A., Amaral, L., Glass, L., Hausdorff, J., Ivanov, P. C., Mark, R., ... & Stanley, H. E. (2000). PhysioBank, PhysioToolkit, and PhysioNet: Components of a new research resource for complex physiologic signals. Circulation [Online]. 101 (23), pp. e215–e220. RRID:SCR_007345.  
  
https://zenodo.org/records/15272720  
  
Alistair E W Johnson, David J Stone, Leo A Celi, Tom J Pollard, The MIMIC Code Repository: enabling reproducibility in critical care research, Journal of the American Medical Informatics Association, Volume 25, Issue 1, January 2018, Pages 32–39, https://doi.org/10.1093/jamia/ocx084  
  